---
title: "Class08"
author: "Chen"
format: html
editor: visual
---

```{r}
wisc.df <- read.csv("/Users/showwhale/Desktop/HsiaoinMac/HsiaoinMac/UCSD/Freshman/Winter/Bioinformatics/class08/WisconsinCancer.csv", row.names = 1)
```


>Q .H How many obervations/samples/patients/rows?

There are 'r nrow(wisc.df) individuals in this dataset.

```{r}
nrow(wisc.df)
```

>Q What is in the '$diagnosis' column? How many of each type?

```{r}
table(wisc.df$diagnosis)
sum(wisc.df$diagnosis == "M")
sum(wisc.df$diagnosis == "B")
```


>Q How many variables/features in the data are suffixed with _mean?

```{r}
cname <- colnames(wisc.df)
cname
cnum <- grep("_mean", cname, value = TRUE)
length(cnum)
```

>Q How many variables/features in total?

```{r}
ncol(wisc.df)
```

>Q Save the diagonises for later reference

```{r}
diagonsis <- as.factor(wisc.df$diagnosis)
```


and remove or exclude this colimn from any of our further analysis
```{r}
wisc.df <- wisc.df[,-1]
```

Let's try clustering this data!

```{r}
wisc.hc <- hclust(dist(wisc.df))
plot(wisc.hc)
```


#Principal Component Analysis

Let's try PCA on this data. Before doing any anlysis like this we should check if our data is well-scaled. Normalizing these data can standardize the variation

```{r}
apply(mtcars, 2, mean)
```
```{r}
apply(mtcars, 2, sd)
```

Let's try a PCA on this car dataset

```{r}
pc <- prcomp(mtcars)
summary(pc)
```
```{r}
biplot(pc)
```
Do scalling first!

```{r}
pc.scale <-prcomp(mtcars, scale = TRUE)
summary(pc.scale)
biplot(pc.scale)
```


## Back to our cancer data set##
Do we need to scale this data set? Yes. Because the spread is very different in each variables.

```{r}
wisc.pr <- prcomp(wisc.df, scale = TRUE)
summary(wisc.pr)
biplot(wisc.pr)
```


Our main PC score plot (a.k.a PC plot, PC1 vs PC2, orientation)

```{r}
attributes(wisc.pr)
```

We need to build our own plot because wisc.pr is too crowded and non informative.

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col = diagonsis)
```



